3D1D modeling of the convective-reactive mixing in RAWDs

### 2021-05-01 paper_mnras/paper.tex
add ulem package
updates
clean template

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
% macros. please check here before defining something new.
\input{../paper_resources/macros/code}
\input{../paper_resources/macros/derivatives}
\input{../paper_resources/macros/nuclides}
\input{../paper_resources/macros/concepts}
\input{../paper_resources/macros/units}
\input{../paper_resources/macros/vectors}
\input{../paper_resources/macros/formatting}
\input{../paper_resources/macros/symbols}

@@

% for author's comments; choose your own color and replace 'fh' and
% 'FH' with your name and initials
\usepackage[usenames,dvipsnames]{color}
\usepackage[]{color}
\newcommand{\fhcom}[1]{{\color{PineGreen}[\emph{FH: #1}]}}
\newcommand{\fhtxt}[1]{{\color{PineGreen}[FH: #1]}}

@@

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[Short title, max. 45 characters]{MNRAS \LaTeXe\ template -- title goes here}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[K. T. Smith et al.]{
Keith T. Smith,$^{1}$\thanks{E-mail: mn@ras.org.uk (KTS)}
A. N. Other,$^{2,\dagger}$
Third Author$^{2,3}$
and Fourth Author$^{3}$
\\
% List of institutions
$^{1}$Royal Astronomical Society, Burlington House, Piccadilly, London W1J 0BQ, UK\\
$^{2}$Department, Institution, Street Address, City Postal Code, Country\\
$^{3}$Another Department, Different Institution, Street Address, City Postal Code, Country\\
$^\dagger$NuGrid collaboration}

@@

% Enter the current year, for the copyright statements etc.
\pubyear{2015}

@@

% Abstract of the paper
\begin{abstract}
This is a simple template for authors to write new MNRAS papers.
The abstract should briefly describe the aims, methods, and main results of the paper.
It should be a single paragraph not more than 250 words (200 words for Letters).
No references should appear in the abstract.
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
keyword1 -- keyword2 -- keyword3
\end{keywords}

@@

This is a simple template for authors to write new MNRAS papers.
See \texttt{mnras\_sample.tex} for a more complex example, and \texttt{mnras\_guide.tex}
for a full user guide.

All papers should start with an Introduction section, which sets the work
in context, cites relevant earlier studies in the field by \citet{1996ApJ...456..902R},
and describes the problem the authors aim to solve \citep[e.g.][]{cyburt:10}.

\section{Methods, Observations, Simulations etc.}

Normally the next section describes the techniques the authors used.
It is frequently split into subsections, such as Section~\ref{sec:maths} below.

\subsection{Maths}
\label{sec:maths} % used for referring to this section from elsewhere

Simple mathematics can be inserted into the flow of the text e.g. $2\times3=6$
or $v=220$\,km\,s$^{-1}$, but more complicated expressions should be entered
as a numbered equation:

\begin{equation}
    x=\frac{-b\pm\sqrt{b^2-4ac}}{2a}.
	\label{eq:quadratic}
\end{equation}

Refer back to them as e.g. equation~(\ref{eq:quadratic}).

\subsection{Figures and tables}

Figures and tables should be placed at logical positions in the text. Don't
worry about the exact layout, which will be handled by the publishers.

An example figure can be seen in \Fig{mass-N14}.
, and tables as
e.g. \Tab{example_table}.

% Example figure
\begin{figure}
  % To include a figure from a file named example.*
  % Allowable file formats are eps or ps if compiling using latex
  % or pdf, png, jpg if compiling using pdflatex

  %\includegraphics[width=\columnwidth]{figs/figure_template/mass-N14}
  \caption{This is an example figure. Captions appear below each
    figure.  Give enough detail for the reader to understand what
    they're looking at, but leave detailed discussion to the main body
    of the text.  This particular figure file will appear when you
    exectute the example notebook in
    \texttt{figs/figure\_template}. Then uncomment the
    \texttt{includegraphics} command in the latex source file.} 
    \lFig{mass-N14}
\end{figure}

Another results is also good \Fig{mass-O16}.
% Example table
\begin{table}
	\centering
	\caption{This is an example table. Captions appear above each table.
	Remember to define the quantities, symbols and units used.}
	\lTab{example_table}
	\begin{tabular}{lccr} % four columns, alignment for each
		\hline
		A & B & C & D\\
		\hline
		1 & 2 & 3 & 4\\
		2 & 4 & 6 & 8\\
		3 & 5 & 7 & 9\\
		\hline
	\end{tabular}
\end{table}

% Example figure
\begin{figure}
	% To include a figure from a file named example.*
	% Allowable file formats are eps or ps if compiling using latex
	% or pdf, png, jpg if compiling using pdflatex
	\includegraphics[width=\columnwidth]{figs/figure_template/mass-O16}
    \caption{Another figure.}
    \lFig{mass-O16}
\end{figure}

\section{Conclusions}

The last numbered section should briefly summarise what has been done, and describe
the final conclusions which the authors draw from their work.

>>>>

% Only include extra packages if you really need them. Common packages are:
\usepackage{rotating}
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{booktabs}
\usepackage{natbib}
% \usepackage[dvipsnames]{xcolor}
\usepackage{textcomp}
\usepackage{xfrac}
\usepackage{afterpage}
\usepackage[normalem]{ulem}
\usepackage{threeparttable} % modify table

% macros. please check here before defining something new.
\input{../paper_resources/macros/code}
\input{../paper_resources/macros/derivatives}
\input{../paper_resources/macros/nuclides}
\input{../paper_resources/macros/concepts}
\input{../paper_resources/macros/units}
\input{../paper_resources/macros/vectors}
\input{../paper_resources/macros/formatting}
\input{../paper_resources/macros/symbols}

@@

% for author's comments; choose your own color and replace 'fh' and
% 'FH' with your name and initials
\usepackage[usenames,dvipsnames]{color}
\usepackage[]{color}
\newcommand{\fhcom}[1]{{\color{PineGreen}[\emph{FH: #1}]}}
\newcommand{\fhtxt}[1]{{\color{PineGreen}[FH: #1]}}
\newcommand{\racom}[1]{{\color{Brown}[\emph{RA: #1}]}}
\newcommand{\ratxt}[1]{{\color{Brown}[RA: #1]}}
\newcommand{\dscom}[1]{{\color{Red}[\emph{DS: #1}]}}
\newcommand{\dstxt}[1]{{\color{Red}[DS: #1]}}

% new commands
\newcommand{\cldfluid}{fluid above the convection zone}
\newcommand{\airfluid}{convective fluid}
\newcommand{\ppmstar}{\code{PPMstar}}

@@

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[3D1D modeling of the convective-reactive mixing in RAWDs]{3D1D
  modeling of the convective-reactive mixing in RAWDs} \title[3D1D
  modeling of the convective-reactive mixing in RAWDs]{3D1D
  hydro-nucleosynthesis simulations. I. Advective-reactive
  post-processing method and its application to H ingestion into He-shell
  flash convection in rapidly accreting white dwarfs}
% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[D. Stephens et al.]{
David Stephens,$^{1}$
Falk Herwig,$^{1,2,\dagger}$\thanks{E-mail: fherwig@uvic.ca}
Paul Woodward,$^{2,3}$
Pavel Denissenkov,$^{1,2,\dagger}$
\newauthor
 Robert Andrassy$^{4,1}$ 
and Huaqing Mao$^{3}$\\
% List of institutions
%\author[0000-0001-6120-3264]{Pavel Denissenkov}
$^{1}$Department of Physics and Astronomy, University of Victoria, Victoria, BC, V8W 2Y2, Canada\\
$^{2}$Joint Institute for Nuclear Astrophysics - Center for the Evolution of the Elements, USA\\
%\author[0000-0001-8087-9278]{Falk Herwig}
$^{3}$LCSE and Department of Astronomy, University of Minnesota, Minneapolis, MN 55455, USA\\
$^{4}$Heidelberg Institute for Theoretical Studies, Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg, Germany\\
$^\dagger$NuGrid Collaboration, \href{http://nugridstars.org}{http://nugridstars.org}\\}

@@

% Enter the current year, for the copyright statement etc.
\pubyear{2020}

@@

% Abstract of the paper
\begin{abstract}
We present two mixing models for post-processing of 3D hydrodynamic
simulations applied to convective-reactive \ipr\ nucleosynthesis in a
rapidly accreting white dwarf (RAWD) with $\mathrm{[Fe/H]} = -2.6$, in
which H is ingested into a convective He shell during a He flash. A 1D
advective two-stream model is formulated with physically motivated
radial and horizontal mixing coefficients constrained by 3D
hydrodynamic simulations. A more traditional approach uses diffusion
coefficients calculated from the same simulations. All 3D simulations
include the energy feedback of the
\nuclei{12}{C}$(\pt,\gamma)$\nuclei{13}{N} reaction from the
entrainment of stably stratified H. Global oscillations of shell H
ingestion in two of the RAWD simulations cause bursts of
entrainment of H and energy feedback into the flow. With the same
nuclear network as in the 3D simulations, the 1D advective two-stream
model reproduces the rate and location of the H burning within the He
shell closely matching the 3D simulation predictions, as well as
qualitatively displaying the asymmetry of the $X_{\mathrm{H}}$
profiles between the up- and downstream. With a full \ipr\ network
the advective mixing model captures the difference in the n-capture
nucleosynthesis in the up- and downstream. For example,
\nuclei{89}{Kr} and \nuclei{90}{Kr} with half-lives of
\unit{3.18}{\minute} and \unit{32.3}{\second} differ by a factor 2 - 10 in the
two streams. In this particular application the diffusion approach
provides globally the same abundance distribution as the advective
two-stream mixing model. The resulting \ipr\ yields from the diffusive
and advective post-processing models are compared with observations of
the exemplary CEMP-r/s star CS31062-050.

\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
convection - hydrodynamics - turbulence - stars: evolution - stars: interiors - stars: white dwarfs - nuclear reactions, nucleosynthesis, abundances.
\end{keywords}

@@

% Herwig 2011, Kippenhahn/MLT, Davis
The details of mixing in convection zones within the stellar interior
modifies the evolution of stars from the main sequence to the final
white dwarf stage or supernova core collapse. The continual mixing of
H into the convective cores of intermediate and massive stars sets a
timescale for the eventual main-sequence turn off. This can be
extended from convective boundary mixing which introduces additional H
fuel to the core. Through this additional mixing of fuel into
convection zones convective boundary mixing has a cumulative effect on
the nuclear-time scale evolution of individual convection zones
including on the mixture and amount of processed elements, the size of
the convection zones and thereby on subsequent evolutionary phases
\citep[for
  example][]{herwig:00c,Young:2005eg,Denissenkov:2012cu,Battino:2016bna,Davis2019,Wagstaff:2020ce}.

Over the nuclear timescale of the H-burning during the main sequence,
the mixing of chemical species is computed using either an
instantaneous or diffusive mixing approximation. The nuclear burning
timescale is usually significantly longer than the mixing timescale
within the convective core leading to the details of the mixing within
the convective zone being sufficiently well modeled with a space-time
average theory. Such a theory, the mixing length theory
\citep[MLT][]{cox:68}, describes the energy transport and mixing
properties of a convective region using averaged quantities over space
and time.  In advanced stages of stellar evolution the nuclear burning
timescales approach convective timescales, and so the details of the
mixing become increasingly important for their structure
\citep[][]{Herwig2011,Collins:2018cp,Cote:2020bo}.  To quantify this
phenomenon, it is useful to define the Damk\"ohler number,
$\mathrm{Da} = \tau_{\mathrm{mix}} / \tau_{\mathrm{nuclear}}$, which
is the ratio of the mixing and nuclear burning timescales. As this
number becomes larger and closer to 1, the details of the mixing have
a greater impact on the burning.

3D hydrodynamic simulations are becoming increasingly accurate in
capturing the process of entrainment of fuel into a convection zone
\citep{meakin:07b,Mocak:2011jk,Woodward:2013uf} as well as the
hydrodynamic feedback resulting from the dynamic nuclear burning of
the ingested fuel
\citep{meakin:06a,Herwig:2014cx,Muller:2016kta,Yoshida:2019be,Andrassy:19,Yadav:2020gh}. Such
simulations predict how species are mixed into the convection zone and
can determine the convective-reactive nucleosynthesis for a limited
number of species without needing a post-processing.  However,
modeling any complicated nucleosynthesis, for example the
\iprn\ \citep{herwig:11}, or the detailed nucleosynthesis in the
merger of a O and a C convection shell \citep{Ritter2018} may require
networks with hundreds or thousands of species that can interact with
each other, which is well beyond the capabilities of any 3D
hydrodynamic code running on modern computing clusters. Therefore, 1D
mixing models are still required to determine any complicated
nucleosynthesis while 3D hydrodynamic simulations can be used to
determine the mixing properties of the convection zone and how the
hydrodynamic instabilities lead to the entrainment of stably
stratified material. For example, the 3D hydrodynamic simulations of
\citet{Ritter2018} modeled a convective O shell with a stable C shell
from a $15\,M_{\odot}$ stellar model of \citet{RitterSE} to obtain
estimates for the expected entrainment rate of the C-rich
material. Using diffusive mixing constrained by the mixing properties
of the convection zone and appropriate entrainment rates, the 1D
large-network nucleosynthetic post-processing models produced
significant amounts of odd-Z elements like P, Cl, K and Sc, which
could explain the underproduction of these elements in current GCE
models~\citep{Ritter2018}.  The nuclear network used in the 3D
simulations only included the energy generation from the
\nuclei{12}{C}$(\nuclei{12}{C},\alpha)$\nuclei{20}{Ne} and subsequent
\nuclei{16}{O}$(\alpha,\gamma)$\nuclei{20}{Ne} reaction.

% 
An interesting case of convective-reactive nucleosynthesis occurs when
H is ingested into a He burning convective shell.  This triggers the
\nuclei{12}{C}$(\pt,\gamma)$\nuclei{13}{N} reaction and, after the
\nuclei{13}{N} beta decays to \nuclei{13}{C}, the
\nuclei{13}{C}$(\alpha,n)$\nuclei{16}{O} reaction can release neutrons
if the temperatures within the He shell are high enough. Sakurai's
object (V4334 Sagittarii) has a unique surface chemical composition
\citep{asplund:99a} that can be explained with H being ingested into
the He-shell flash convection zone of this post-AGB
star~\citep{herwig:11}. This results in the above chain of reactions
and produces neutron densities high enough to be in the \ipr~regime
($N_{n} \approx 10^{12} - 10^{16}$\,cm$^{-3}$) \citep{cowan:77}. The
energy generation from this H ingestion can be significant enough to
cause a split of the He convective zone in 1D stellar evolution
models. To study the nucleosynthesis in Sakurai's object,
\citet{herwig:11} used a spherically symmetric diffusive mixing model
based on MLT convective velocities, even though in those conditions
several MLT assumptions are not satisfied. This is shown explicitly in the
3D hydrodynamic simulations of this very energetic event which bring
about a global oscillation of shell H ingestion, that causes
large-scale, non-radial and fast flows \citep{Herwig:2014cx}.

A more quasi-static convective-reactive case of H-ingestion into a He
shell has been found in the models of rapidly accreting white dwarfs
(RAWDs) by \cite{Denissenkov:2017ba}. The 3D hydrodynamic simulations
of \citet{Denissenkov:19} quantified the H-entrainment rates, however
the diffusive mixing used in the post-processing was taken directly
from the 1D stellar evolution models. The neutron densities in RAWDs
reach $N_{n} \approx 10^{14}$\,cm$^{-3}$ resulting in
\ipr~nucleosynthesis.  The RAWD heavy element production could result
in a significant contribution of Kr, Rb, Sr, Y, Zr, Nb and Mo to the
solar composition \citep{Cote2018}. The convective-reactive flows in
RAWDs are fed by the \nuclei{12}{C}$(\pt,\gamma)$\nuclei{13}{N} energy
generation that comprises only $2-3\%$ of the total luminosity within
the He shell which does not lead to a global oscillation of shell H
ingestion \citep{Denissenkov:19}.

A solution to the concerns about the validity of diffusive mixing in a
convective-reactive environment and the inability of 3D hydrodynamic
simulations to simultaneously perform complex nucleosynthesis
computations is to use a 1D advective mixing model.  One such model
has been adopted for use in the post-processing nucleosynthesis of
stellar evolution models from the Monash group, \code{MONSOON}
\citep{cannon1993, Henkel2017}. The model contains two adjacent
streams of fluid flow \citep[see our \Fig{elevator} and Fig.~1
  in][]{Henkel2017}, one with fluid moving upwards and another with
fluid moving downwards. The two streams have an enforced horizontal
mixing in order to conserve mass but it can also add in additional
horizontal mixing. The radial transport velocities as well as the
additional horizontal mixing within a convection zone are estimated
with MLT.

A limitation in the two-stream methodology of \code{MONSOON} arises in
the treatment of the additional horizontal mixing. The strength of
this mixing has no dependence on the structure of the flow within the
convection zone. Is the mixing between the two streams that represent
the dominant dipolar flows of core convection the same as if the flow
field was at smaller angular scales like in shell
convection~\citep{chandrabook1961}? There is a dependence on the
horizontal mixing based on where a cell is within the convection zone
but shouldn't the mixing between the two streams, in low \mach~number
flows, be stronger near the convective boundaries where the fluid is
forced to overturn well before the convective
boundaries~\citep{Jones:2017kc}? These limitations are addressed in
the advective mixing model of this work.

In this paper, we describe our adaptation of an advective two-stream
model for post-processing of detailed 3D hydrodynamic simulations. We
create 3D hydrodynamic simulations of a RAWD model
from~\citet{Denissenkov:19} to quantify the mixing of H into the He
shell and to simulate the global flow including the energy feedback
from nuclear burning of entrained material. After extracting the
mixing information from the 3D hydrodynamic simulations for the 1D
diffusive ~\citep{Jones:2017kc} and advective mixing models, the time
evolution of the H burning in both of them is calculated.  The details
of the 3D hydrodynamic simulations, as well as the diffusive and
advective mixing approaches are outlined in \Section{allmethods}.
\Section{results} discusses the flow properties of the simulations and shows the post-processing of the 3D hydrodynamic
simulations with the diffusive and advective mixing routines.
\Section{conclusions} describes the implications of such results and
further applications.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Methods}
\label{sec:allmethods}

\subsection{PPMstar simulations}
\label{sec:ppmsim}

% Table of runs
\begin{table*}
	\centering
	\caption{Summary of the \ppmstar{} simulations that were computed for this
          work. The entrainment rates are the slopes of the linear fits shown in
          \Fig{entrain}, while the different definitions of the convective boundary are
          discussed in detail in \Sections{convbound}{modelConstraints}.}
	\label{tab:ppmstarmodels}
  \begin{threeparttable}
	\begin{tabular}{ccccccccc} % four columns, alignment for each
		\toprule
		Run ID & Grid & $t_{\mathrm{sim}} \unitspace (\mathrm{min})$ & $\langle \tau_{\mathrm{conv}}^{i} \unitspace (\mathrm{min}) \rangle$ & $L_{\mathrm{He}} \unitspace (L_{\odot})$ & $r_{\mathrm{b}, \mathrm{SC}}^{ii} \unitspace (\mathrm{Mm})$ & $r_{\mathrm{b}, v_{\perp}}^{iii} \unitspace (\mathrm{Mm})$ & $\sigma_{r_{\mathrm{b}}, v_{\perp}} \unitspace (\mathrm{Mm})$ & $\dot{M}_{\mathrm{e}}^{iv} \unitspace (M_{\odot} \unitspace \mathrm{s}^{-1})$\\
		\midrule
		N15 & $768^{3}$ & 1634 & 19 & $1.46 \times 10^{8}$ & $23.86$ & $23.43$ & $0.48$ & $1.07 \times 10^{-11}$  \\
		N16 & $1536^{3}$ & 744 & 18 & $1.46 \times 10^{8}$ & $23.84$ & $23.56$ & $0.46$ & $7.21 \times 10^{-12}$  \\
		N17 & $1152^{3}$ & 631 & 9 & $14.6 \times 10^{8}$ & $26.74$ & $26.69$ & $1.19$ & $1.08 \times 10^{-10}$  \\
		\bottomrule
	\end{tabular}
  \begin{tablenotes}
    \small
  \item \textit{Notes}: $^{i}$ The average convective turn over time during the
    quasi-static phase of each run ($46 < t < 300 \unitspace \mathrm{min}$); $^{ii}$ The
    initial Schwarzschild boundary as followed in the Lagrangian
    coordinates at $t = \unit{299}{min}$; $^{iii}$ Boundary where $\partial v_{\perp} / \partial r$ has a
    minimum at $t = \unit{299}{min}$; $^{iv}$ Entrainment rate of the H-rich 
    \cldfluid{} 
  \end{tablenotes}
  \end{threeparttable}

\end{table*}

The advective and diffusive post-processing methods introduced here are applied
to 3D hydrodynamic simulations of He-shell flash convection in a rapidly
accreting white dwarf \citep{Denissenkov:2017ba}. The initial stratification has
been taken from the stellar evolution model G with the metallicity [Fe/H]\,$= -2.6$
from \cite{Denissenkov:19}.

As in previous work \citep{Herwig:2014cx,Jones:2017kc} we use the \ppmstar{}
code of \citet{woodward15} with additional details provided by
\citet{Andrassy:19}. The explicit Cartesian grid code is based on the
Piecewise-Parabolic Method \citep[PPM;][]{woodward_colella81,
woodward_colella84, colella_woodward84, woodward86, woodward07}, and tracks the
advection of concentrations in a two-fluid scheme using the Piecewise-Parabolic
Boltzmann method \citep[PPB;][]{woodward86,woodward15}.

The luminosity from the \nuclei{4}{He} burning within the convection zone is
modeled with a constant volume heating. The entrained H reacts with the abundant
\nuclei{12}{C} from the triple-$\alpha$ via the
\nuclei{12}{C}$(\pt,\gamma)$\nuclei{13}{N} reaction. This reaction rate is
computed using the analytic form of the rate from \citet{angulo:99} with no
screening factor. We ignore the subsequent beta decay of \nuclei{13}{N} leaving
a total energy release per reaction of $Q = \unit{1.943}{MeV}$. The stably
stratified \cldfluid{} contains 89.4\% by
number of H and the \airfluid{} contains 14.3\% by
number of \nuclei{12}{C}. The \ppmstar{} simulations done in this work are summarized
in \Tab{ppmstarmodels}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Diffusive mixing model}
\label{sec:Dmethod}

\citet{Jones:2017kc} inverted the diffusion equation to derive a
radius-dependent diffusion coefficient, which produces in 1D the same
redistribution of species over the time frame of analysis as that given by their
spherically-averaged 3D simulations of O-shell convection. They measured the
rate of change in the radius-dependent mass fraction $X(r,\ t)$ of a species by
computing the difference between the mass fraction profiles at two different
points in time with some time averaging applied around them. With $\partial X /
\partial r$ also known from the spherical averages, \citet{Jones:2017kc} solve
for the unknown diffusion coefficient, which we call $D_{\mathrm{FV}}(r)$, in
the 1D Eulerian diffusion equation
%
\begin{equation*}
\frac{\partial X}{\partial t} = \frac{\partial}{\partial x}\left( D_{\mathrm{FV}}(r)
\frac{\partial X(r)}{\partial x} \right),
\end{equation*}
%
where they set $x = r$. We have improved upon this method by mapping
the results of the input 3D Eulerian simulations to a mass coordinate
$m(r)$ and inverting the Lagrangian diffusion equation
%
\begin{equation}
\frac{\partial X}{\partial t} = \frac{\partial}{\partial m}\left( \sigma(m)
\frac{\partial X(m)}{\partial m} \right) + \frac{\dot{q}}{\rho},
\label{eq:diff-inv-burn}
\end{equation}
%
where $\sigma = (4\pi r^2 \rho)^2 D_\mathrm{FV}(r)$ is the Lagrangian diffusion
coefficient, and $\dot{q}(m,\ t)$ is the destruction rate of the species by
nuclear reactions. This new approach has the following advantages. It removes
the effect of thermal expansion and contraction from $\partial X / \partial t$.
This effect has negligible influence on the results of~\citet{Jones:2017kc}, but
it is essential to account for in the expanding convection zone in these RAWD
simulations. It also properly takes into account the spherical geometry of the
problem with the radial dependence of the density and species. Included in our
RAWD simulations but not in the O-shell convection simulations of
\citet{Jones:2017kc} are nuclear reactions between species of the two distinct
fluids containing H and \nuclei{12}{C}.

In order to invert \eq{diff-inv-burn} we need to take the difference of two
averages of the fractional volume profiles, or -- in this case -- the mass
fraction profile $X(r,\ t)$. For that we estimate the time in which it takes to
diffuse across one mixing length:
%
\begin{equation*}
\Delta t = \frac{(\alpha_\mathrm{MLT} H_\mathrm{P})^2}{4D_{\mathrm{FV}}}
\end{equation*}
%
where we take $H_\mathrm{P}$ and $D_{\mathrm{FV}}$ at the radius where
$D_{\mathrm{FV}}$ is at its maximum. This diffusive timescale is $\Delta t
\approx \unit{2.5}{\minute}$.

\ppmstar{} simulations output a \textit{dump} every several thousand time steps.
For example in N16 every $\approx 2121$ time steps, where each time step is
$\approx \unit{\natlog{1.3}{-2}}{\second}$, a single \textit{dump} is output.
For each \textit{dump} at time $t_\mathrm{D}$, $D_\mathrm{FV}(r)$ is determined
based on two profiles, $X(r,\ t_\mathrm{D}\pm \Delta t)$, each being averaged
over $\Delta t$.

Despite taking the burning term into account in \eq{diff-inv-burn}, the profile
of the entrained material cannot be completely recovered, and thus the method
cannot provide a diffusion coefficient in the lower part of the convection zone
where all of the entrained material is burned (top panel \Fig{DvrDVF-compare}).
%
\begin{figure}
  \centering
	\includegraphics[width=\columnwidth]{figs/N16-DHp-compare-dump650.pdf}
	\includegraphics[width=\columnwidth]{figs/D-vr-FV-compare.pdf}
  \caption{Top: $D_\mathrm{vr}$ according to \eq{dvr}, and
          $D_\mathrm{FV}$ according to the direct determination from the
          $X_\mathrm{H}$ profiles as well as the pressure scale
          height at $t = \unit{299}{min}$. Bottom: $D_\mathrm{vr}$ (grey lines) and the composite
          $D_\mathrm{FV+vr}$ (colour lines) according to the direct
          determination from the $X_\mathrm{H}$ at various times, merged
          with $D_\mathrm{vr}$ in the lower part of the convection zone (see
          text for details). }
  % Top Fig made in ../notebooks/Diffusion/Lagrangian_diffusion_analysis-make-shell-files.ipynb
  % Bot Fig made in ../notebooks/Diffusion/D-vr-FV-compare.ipynb
  \label{fig:DvrDVF-compare}
\end{figure}
%
Therefore we also determine the diffusion coefficient from the
spherically averaged radial velocity $v_\mathrm{PPM}$ from the 3D
hydrodynamic simulation using the correction for the near-boundary
regions recommended by \citet{Jones:2017kc} based on O-shell
simulations
%
\begin{equation}
D_\mathrm{vr} = \frac{1}{3} v_\mathrm{PPM}\times \min \left(l,|r-r_0|\right),
\label{eq:dvr}
\end{equation}
%
where $v_\mathrm{PPM}$ is the rms of the radial velocity on spherical shells
from the 3D \ppmstar\ simulation of the He-shell convection, $l =
\alpha_{\mathrm{MLT}} H_P$ is the mixing length, and $|r-r_0|$ is the distance
to the He-shell or another convective boundary that is located at the radius $r_0$. For
$\alpha_{\mathrm{MLT}} = 1.6$, this prescription provides the best fit to the
diffusion coefficient derived from the hydrodynamic simulations of O-shell
convection simulations \citep[see Figure 21 in][]{Jones:2017kc}. However that
diffusion coefficient recommendation was focused on the behaviour of mixing near
and across the top convective boundary (CB), and no attempt was made to model
the discrepancy between the diffusion coefficient determined from the
spherically averaged 3D hydrodynamic abundance profile evolution and the
diffusion coefficient determined according to \eq{dvr}. This is clearly seen at
the middle of the convection zone, $r = \unit{5.5}{Mm}$, in Figure~21 of
\citet{Jones:2017kc}. In the RAWD case the burning of entrained H-rich material
takes place approximately in the middle of the convection zone where this
difference between $D_\mathrm{vr}$ and $D_\mathrm{FV}$ is largest ($\approx 0.4
\dots \unit{1.2}{dex}$, \Fig{DvrDVF-compare}).

Where we are able to determine $D_\mathrm{FV}$ it is the more accurate diffusion
coefficient that describes the underlying evolution of the spherically averaged
3D hydro profiles of these simulations better, for obvious reasons. In order to be able to use this
more accurate diffusion coefficient we combine it in the lower half with the
values of $D_\mathrm{vr}$ according to \eq{dvr}. We determine the radius
$r_\mathrm{Dmax}$ where $D_\mathrm{FV}$ has its maximum. Then
\begin{equation*}
  D_\mathrm{FV+vr}= \begin{cases}
  D_\mathrm{FV}, & \text{if } r > r_\mathrm{Dmax} , \\
  max(D_\mathrm{FV},D_\mathrm{vr}), & \text{otherwise}.
\end{cases}
\end{equation*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Advective mixing model}
\label{sec:advectmodel}

The 1D advection is formulated using a two-stream approach in which
one stream transports species radially upwards, while the other stream
transports them radially
downwards~\citep[][\Fig{elevator}]{cannon1993, Henkel2017}. Our
adaptation of this advective two-stream (ATS) model to post-processing
of 3D hydrodynamic simulations constitutes a reduced-dimensionality
1.5D advection model.

By default, the two streams are taken to be equal in their surface
areas which is justified by the analysis of the 3D flow properties
(see \Section{flow}) and the discussion in
\Section{radialMix}. The model can have horizontal mixing between up- and
downstream cells that are adjacent to each other.

There are $N$ cells per stream. The index $i$ refers to the spatial index of the
$2N$ cells and is used where the equations are agnostic to whether the cell is
within an up- or downstream. For computational and numerical simplicity, the
indexing is done such that the downstream is inverted and \textit{stitched} onto
the top of the upstream. The indexing starts at the bottom of the upstream at
$i=1$ and so cell $i=N+1$ is at the top of the downstream. The index $k$ refers
to species.

The discretization distinguishes variables defined on the cell boundaries which
have spatial half-integer indices, $i+1/2$, while those that represent cell
averages have spatial integer indices $i$.

\begin{figure}

  \includegraphics[width=0.5\columnwidth]{figs/Advection-Streams.pdf}
  \centering
  \caption{An illustration of the two-stream model including all of the fluxes
          for a cell within the convection zone; $\alpha$ the radial mass flux
          (\Section{radialMix}), $\beta$ the enforced horizontal mass flux and $\gamma$
          the additional horizontal mass flux (\Section{horizontalMix}). The index $i$
          refers to the upstream cell which has an adjacent downstream cell $2N + 1 - i$.}
  % Fig made in ../notebooks/Methods/Advection-Streams.xcf
  \label{fig:elevator}

\end{figure}

\subsubsection{Discretized equations}
\label{sec:discretized}

To formulate the equations we start off with the conservation of mass equation
and then apply the divergence theorem to it to yield
%
\begin{equation}
  \label{eq:masscons}
  \partial_{t} \int_{V} \rho \mathrm{d}V = - \oint_{\partial V} \rho \vec{v} \cdot \mathrm{d}\vec{a}
\end{equation}
%
The term on the right hand side is interpreted as a sum of mass fluxes through
the boundary of the given volume, $V$. The term on the left hand side is the rate of
change of the mass contained within that volume, $V$. This advective model conserves
the total mass of every cell which requires that the sum of all mass fluxes at
every cell is zero. Recasting \eq{masscons} into the partial densities of every
species, integrating both sides and applying the constraint that the mass of
every cell is constant at all times leads to
%
\begin{equation}
  \label{eq:fluxes0}
  \partial_{t} \left( \sum_{k} m_{k,i} \right) = - (\sum_{k}\sum_{j=1}^{5} F_{k,i}^{j}) = 0
\end{equation}
%
where the index $j$ refers to a specific species mass flux on the
surface of cell $i$. The mass, $m_{k,i}$, of a given species within a
cell can change however the total mass within that cell cannot. This
is implicitly satisfied with \eq{fluxes0}. Therefore these equations
only transport the mass of a given species throughout the convection
zone. \Eq{fluxes0} includes all fluxes in the advective model and it
is integrated explicitly.

\subsubsection{Radial mixing and boundary conditions}
\label{sec:radialMix}

Within the convection zone, the bulk transport of species is through the radial
direction. The radial mass flux, $\alpha$, on each cell's boundary is given by
%
\begin{equation}
  \label{eq:alpha}
  \alpha_{i+1/2} = 2 \pi r^{2}_{i+1/2} \rho_{i+1/2} v_{i+1/2}
\end{equation}
%
where $r$, $\rho$, $v$ are the radius, density and velocity defined on
a cells boundary. The velocity is a positive-definite quantity in this
model; the species mass fluxes will explicitly carry the appropriate
sign for transport. In order for this model to remain consistent with
the nearly-hydrostatic equilibrium of the underlying 3D stellar
hydrodynamic simulation, the net mass flux at every radius must be
zero to ensure that there is no net mass transport in the radial
direction. This means that
%
\begin{equation}
  \label{eq:fluxlayer}
  \alpha_{i+1/2} = \alpha_{2N+1/2-i}
\end{equation}
%
